---
title: "Assignment"
author: "Malvin"
date: "18/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE,message=FALSE}
library(tidyverse)
library(ggplot2)
```

# Question 1
## 1.1

```{r}
a <- 5.0
c <- 10
t <- 12
B <- a+c/t
```
The answer for b = `r B`


```{r}
vec_1 <- c(44, 5, 0, "k")
vec_2 <- c(1, 2, 3, FALSE)
vec_3 <- c("a", "b", "c", TRUE)
vec_4 <- c(1, 2, 3, "4")

a<-typeof(vec_1)
b<-typeof(vec_2)
c<-typeof(vec_3)
d<-typeof(vec_4)
```
the first vector is a `r a` vector

the second vector is a `r b` vector

the third vector is a `r c` vector

the last vector is a `r d` vector

# 1.2 conditional Subsetting
### 1.2.1
```{r}
vec <- c(141.2741, 141.8392, 140.5730,NA, 141.3571, 139.8689, 
         137.9384, NA,139.6620, 138.2521, 138.8635, 141.3885, NA)

vec_1<-c(na.omit(vec,na.rm=TRUE))
```
the new vector list is `r vec_1`

I used na.omit to remove all missing values to the "vec" vector that was given

### 1.2.2

```{r}
vec_2 <-c(subset(vec,is.na(vec)))
```
the new vector of missing values is `r vec_2`

I subset the vector into only giving me a list with missing values


### 1.2.3

```{r}
len<-length(vec_2)
```
There are `r len` missing values in vec

### 1.2.4

```{r}
student<-c("First Student","Second Student","Third Student")
name<-c("Malvin","Rotondwa","Heavy")
surname<-c("Maposa","Jacobs","Petla")
age<-c(19,25,46)
gender<-c("Male","Female","Male")
df<-data.frame(student,name,surname,age,gender)
print(df)
```


# Question 2:Functions in R

```{r}
t<-c(1,2,3,4,5)
mean_function<-function(t){
  z<-mean(t,na.rm=TRUE)
  return(z)
}
mean_function(t)
```
the mean is `r t`












# Question 3
### 3.1

```{r}
g <- c("MALE","FEMALE","FEMALE","MALE","FEMALE","MALE","MALE","FEMALE", "MALE","MALE","FEMALE","MALE","MALE",
"FEMALE","MALE","MALE","MALE")
ifelse(g=="MALE",1,2)
```
### 3.2.1

```{r}
x <- rnorm(1000, 120, 5)
y <- sample(c("M","F"),size = 1000, replace = TRUE)
df<-cbind(x,y)
```

### 3.2.2

```{r}
df <- data.frame(cbind(x,y))
```

### 3.2.3


```{r}
df %>% group_by(y) %>% count()
```

### 3.2.4
```{r}
df %>% group_by(y) %>% count() %>% 
  ggplot(aes(x=y,y=n,fill=y))+
  geom_col()+
  labs(title="Total number of Males and Memales",x="Gender",y="total number of males and females")
```


### 3.2.5

```{r}
df$x<-as.double(df$x)
df<-df %>% group_by(y) %>% 
  mutate(total=sum(x))
print(df)
```
```{r}
df %>%
  ggplot(aes(x=y,y=total,fill=y))+
  geom_col()
```


